---
title: "Understanding what Influences House Prices"
author: "me"
date: "11/5/2020"
output: html_document
---

<style type="text/css">
h1.title {
font-size: 40px;
text-align: center;
}
h4.author {
font-size: 40px;
text-align: center;
}
</style>

```{r setup, include=FALSE}
# load packages here
library(fastDummies) # helps create dummy variables
library(corrplot)  # needed for correlation matrix
library(tidyverse)
library(ggfortify)  # plot glmnet objects using ggplot instead of base R
library(car)  # needed for VIFs
library(bestglm)  # for stepwise variable selection methods
library(leaps)  # for stepwise methods
library(gridExtra)  # for laying out multiple ggplots on same plot
library(glmnet)  # for ridge, lasso, and elastic net
set.seed(12345)  # make sure to set your seed when doing cross validation!
sz <- 15  # set size to use when plotting
```

### Background and Introduction






### Methods and Results




The following table displays the variable names in this data set, along with their descriptions.


Variable       | Description
-------------- | -------------
x              | Observation Number
LotArea        | Size of Lot in square feet
YearBuilt      | Year the house was originally built in
YearRemodADD   | Year of the most recent Remodal or Addition, if none originally build date
TotalSF        | Total number of square feet within the house
FullBaths      | Total number of full bathrooms in the house
HalfBaths      | Total number of half baths in the house
TotRmsAbvGrd   | Number of rooms that are Ground level and above
SaleType*      | Describes the condition of the house at time of sale and type of transaction
LotConfig*     | Describes how the house is situated in relation to other houses and the roads
BldgType*      | Describes the type of house (one family, duplex, etc.)
SalePrice      | The price of the sale in US Dollars

* These are categorical variables and will have dummy variables created in the model. However, to simplify the descriptions, the whole categorical variable is described.


### Exploratory Data Analysis
```{r}
# read in the data set and explore
houses <- read.csv(file = "Project 2 Simplified Dataset.csv")
houses2 <- read.csv(file = "House Prices with Dummies.csv")





# exploratory data analysis



```
#### EDA Results




### Creating the Model
```{r}
# Model Creation

# model summary

```


### Checking assumptions
Linear
```{r}

```
[words]

Independent 
[words]

Normality
```{r}

```
[words]

Equal Variance/Homoscadicity 
```{r}

```
[Words]


No Influential Points
```{r}

```
[Words]

All Predictor Variables Included
[Words]

Multicollinearity
```{r}

```
[words]

Transformation of data
```{r}

```
[words]

### Variable Selection (Maybe in assumptions??)
[words]
```{r}

```
[words]
interactions
```{r}

```
[words]


### Final Model
```{r}

```
[words]

Measurements of Model
```{r}

```


conf int of model
```{r}

```


### Summary and Conclusion

[words]


